Compiling Python Modules to Native Parallel Modules Using Pythran and OpenMP Annotations

### 2013-10-26 doc/papers/sc2013/paper.tex
SC2013: limit line length

Since its birth in December 1989, the Python language~\cite{rossum97} has proved
to be useful in various domains, ranging from system administration to web
services, thanks to its dynamicity, expressiveness, rich ecosystem and
% Mehdi : battery-included ??????
``batteries included'' standard library. It is also widely used in scientific
computing~\cite{Oliphant2007}, thanks to the SciPy~\cite{scipy} project
which provides a set of scientific and numeric tools, ranging from linear
algebra to ordinary differential equation solvers or generic algorithms.

@@

It is now common for scripting language to have C bindings. To take advantage of
compiled code, and to overcome the GIL limitations, Python developers has to write
parallel C/C++ functions and the associated boilerplate based on the Python C
API~\cite{pythoncapi}. Tools have been developed to relieve the user from this
cumbersome task, notably SWIG~\cite{swig2003} that relies on an interface specification supplied by the programmer to generate the glue, or
\texttt{boost::python}~\cite{boostpython2007} that relies on C++ templates and
type overloading facilities to guide translation.

@@

Copperhead~\cite{copperhead2011} is a functional, data parallel language
embedded in Python. It uses $n$-uplets, NumPy arrays and lists as its core data
structure and prohibits usage of many control-flow operators such as loops,
enforcing the use of the \texttt{map}, \texttt{filter} or \texttt{reduce}
intrinsics to exhibit parallelism. But it can be efficiently compiled to either
CUDA or C++ with calls to the Thrust\footnote{\emph{cf}.\
\texttt{http://thrust.github.com/}} library. Python decorators are used to
identify hot-spots that are JIT-compiled to native code.

>>>>

Since its birth in December 1989, the Python language~\cite{rossum97} has proved
to be useful in various domains, ranging from system administration to web
services, thanks to its dynamicity, expressiveness, rich ecosystem and
``batteries included'' standard library. It is also widely used in scientific
computing~\cite{Oliphant2007}, thanks to the SciPy~\cite{scipy} project
which provides a set of scientific and numeric tools, ranging from linear
algebra to ordinary differential equation solvers or generic algorithms.

@@

It is now common for scripting language to have C bindings. To take advantage of
compiled code, and to overcome the GIL limitations, Python developers have to
write parallel C/C++ functions and the associated boilerplate based on the
Python C API~\cite{pythoncapi}. Tools have been developed to relieve the user
from this cumbersome task, notably SWIG~\cite{swig2003} that relies on an
interface specification supplied by the programmer to generate the glue, or
\texttt{boost::python}~\cite{boostpython2007} that relies on C++ templates and
type overloading facilities to guide translation.

@@

Copperhead~\cite{copperhead2011} is a functional, data parallel language
embedded in Python. It uses $n$-uplets, NumPy arrays and lists as its core data
structure and prohibits usage of many control-flow operators such as loops,
enforcing the use of the \texttt{map}, \texttt{filter} or \texttt{reduce}
intrinsics to exhibit parallelism. But it can be efficiently compiled to either
CUDA or C++ with calls to the Thrust\footnote{\emph{cf}.
\texttt{http://thrust.github.com/}} library. Python decorators are used to
identify hot-spots that are JIT-compiled to native code.
